Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-02-10 2015-eamt-melsa/2015-eamt-melsa.tex
some more compressing
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58809 72bbbca6-d526-0410-a7d9-f06f51895060
trimming
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58810 72bbbca6-d526-0410-a7d9-f06f51895060

For creating the test corpora, and for providing a source language corpus 
for training, and a target language corpus for scoring, we used four parallel corpora: 
\begin{itemize}
\item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
  of Breton and French has been collected specifically for
  lexical-selection experiments from translations produced by
  \emph{Ofis ar Brezhoneg} `The Office of the Breton
  language'.\footnote{In 2010 the administrative status was changed
    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
    of the Breton language'. The corpus was published before this date
    so we use the original name.}  It contains some parallel data
  previously described in \cite{tyers09b}, and also some new data. The
  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se/OfisPublik.php}.}
\item \textbf{South-East European Times} (\texttt{SETimes}): Described
  in \cite{tyers10}, this corpus is a multilingual corpus of the
  Balkan languages in the news domain. The Macedonian and English
  portion will be used.
\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
  and Spanish parallel corpus made from the translation memories of
  the \emph{Herri Arduralaritzaren Euskal Erakundea} `Basque Institute
  of Public
  Administration'.\footnote{\url{http://opendata.euskadi.net/w79-contdata/es/contenidos/ds_recursos_linguisticos/memorias_traduccion/es_izo/memorias_traduccion_izo.html}}
\item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
  Described by \cite{koehn05a}, this is a multilingual corpus of the
  European Union official languages. We are using the English--Spanish
  data from version 7.
\end{itemize}

@@

\begin{itemize}
\item \textbf{Linguist-chosen defaults}. 
%! \comment{FSM: This should go to the section explaining Apertium.}  
  A bilingual dictionary in an
  Apertium language pair contains correspondences between lexical
  forms. The dictionaries allow many lexical forms to translate to one
  lexical form. For example, \emph{ahizpa} and \emph{arreba} in Basque
  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
    difference being that \emph{ahizpa} is a sister of a female, and
    \emph{arreba} is a sister of a male.} But a single lexical form
  may not have more than one translation without further
  processing. If there are many possible translations of a lexical
  form, then one must be marked as the \emph{default} translation. For
  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
  and \emph{piso} `flat' in Spanish. If the translation correspondence
  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
  the most general, it is marked as the default by the author of the dictionary. % linguist who is
%  creating the bilingual dictionary.

@@

%The authors thank Filip Petkovski for his help in running some of the
%experiments. They also acknowledge partial support from the Ministry
%of Science and Innovation of the Spanish Government through project
%TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
%Research Grant) coordinated by the Universidad Complutense de Madrid.
%
\bibliographystyle{apalike}
\bibliography{2015-eamt-melsa}

>>>>

For creating the test corpora, and for providing a source language corpus 
for training, and a target language corpus for scoring, we used four parallel corpora: 
\addtolength{\itemsep}{-0.15in}
\begin{itemize}
\item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
  of Breton and French has been collected specifically for
  lexical-selection experiments from translations produced by
  \emph{Ofis ar Brezhoneg} `The Office of the Breton
  language'.%\footnote{In 2010 the administrative status was changed
%    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
%    of the Breton language'. The corpus was published before this date
%    so we use the original name.}  
  It contains some parallel data
  previously described in \cite{tyers09b}, and also some new data. The
  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
\item \textbf{South-East European Times} (\texttt{SETimes}): Described
  in \cite{tyers10}, this corpus is a multilingual corpus of the
  Balkan languages in the news domain. The Macedonian and English
  portion will be used.
\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
  and Spanish parallel corpus made from the translation memories of
  the \emph{Herri Arduralaritzaren Euskal Erakundea} `Basque Institute
  of Public
  Administration'.\footnote{\url{http://opendata.euskadi.net/w79-contdata/es/contenidos/ds_recursos_linguisticos/memorias_traduccion/es_izo/memorias_traduccion_izo.html}}
\item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
  Described by \cite{koehn05a}, this is a multilingual corpus of the
  European Union official languages. We are using the English--Spanish
  data from version 7.
\end{itemize}

@@

\begin{itemize}
\addtolength{\itemsep}{-0.05in}
\item \textbf{Linguist-chosen defaults}. 
%! \comment{FSM: This should go to the section explaining Apertium.}  
  A bilingual dictionary in an
  Apertium language pair contains correspondences between lexical
  forms. The dictionaries allow many lexical forms to translate to one
  lexical form. For example, \emph{ahizpa} and \emph{arreba} in Basque
  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
    difference being that \emph{ahizpa} is a sister of a female, and
    \emph{arreba} is a sister of a male.} But a single lexical form
  may not have more than one translation without further
  processing. If there are many possible translations of a lexical
  form, then one must be marked as the \emph{default} translation. For
  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
  and \emph{piso} `flat' in Spanish. If the translation correspondence
  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
  the most general, it is marked as the default by the author of the dictionary. % linguist who is
%  creating the bilingual dictionary.

@@

%The authors thank Filip Petkovski for his help in running some of the
%experiments. They also acknowledge partial support from the Ministry
%of Science and Innovation of the Spanish Government through project
%TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
%Research Grant) coordinated by the Universidad Complutense de Madrid.
%
\bibliographystyle{apalike}
\small{
\bibliography{2015-eamt-melsa}
}
